Bounded Model Checking for Embedded Systems \\ (Position Paper)

### 2016-01-04 paper.tex
minor changes

Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems, become interesting solutions only if they can reliably
perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.

@@

A feasible alternative to check for properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}. In the former, the goal to check that $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed). The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.

In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}. Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.

Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a \textcolor{red}{conforme conversamos, eh um novo mesmo?} new algorithm for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.

@@

It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior. Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems. For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check for overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. Indeed, DSverifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup. %Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.

@@

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs (ESBMC)~\cite{Cordeiro12}, which was later extended to C++98~\cite{ECBS13}. \textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}. Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC). \textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}. \textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UVA application was tackled, in order to verify its attitude controllers \textcolor{red}{que artigo do iury mencionar?}. Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.

@@

@@

>>>>

Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs) and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may also cost human-lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports also have the potential to compromise patients' health, with serious consequences.

@@

A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}. In the former, the goal is to check whether $\phi$ holds in all states that are reachable from an initial state, within $k$ steps ($k \geq 0$ is assumed). The latter, in turn, aims to check that whenever $\phi$ holds, in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next system state $s_{k+1}$.

In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain violations related to data races~\cite{Donaldson10,Kinductor} and do respect time constraints, which were specified during the system design phase~\cite{EenS03}. Besides, \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}.

Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence, in the associated literature, that model checking of embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15} because of the distinctive characteristics mentioned earlier, which influence the embedded system development and verification process.

@@

It is worth noticing that, currently, SMT-based BMC approaches check code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior. Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if the source code is correctly written, but also if it will properly respond in real environments or under external problems. For instance, the anti-lock braking system software of a vehicle model can be bug free, but it may not work correctly if a sensor is damaged.

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions and, ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

Scratch is also another example of software model checker, which uses knowledge about the system to be verified and the underlying hardware, for detecting races related to direct memory access (DMA) for the Cell BE processor~\cite{Donaldson10}. The tool also uses BMC to detect DMA races, and bounded model checking with k-induction to prove absence of races.

%Finally, DSVerifier has already been used for verifying digital filters and controllers, which even include UAV applications.

@@

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}.

\textbf{(RP2)} This SMT-based BMC approach was further developed to verify deadlock, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering partial-order reduction and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}. Such verification methodologies and respective tools were also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were confirmed by the benchmark's creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC).

\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated, demonstrating is effectiveness in some real-life embedded system applications~\cite{Gadelha15,Rocha15}.

\textbf{(RP4)} The proposed SMT-based context-BMC was also extended to verify overflow, limit cycle, time constraints, stability, and minimum phase in discrete-time systems, in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, in order to verify its attitude controllers~\cite{Bessa16}. Lastly, yet importantly, ESBMC represents one of the most prominent BMC tools (for verifying embedded systems), as observed in the international software verification competitions~\cite{MorseCNF13,MorseRCN014}.

@@

\bibitem{Cordeiro09}
Cordeiro L., Fischer B., Chen H., Marques-Silva J.:
\newblock {\em Semiformal Verification of Embedded Software in Medical Devices Considering Stringent Hardware Constraints.} 
\newblock In: International Conference on Embedded Software and Systems, pp. 396--403, 2009.

@@

\bibitem{Bessa16}
Bessa I., Ismail H., Cordeiro L., Chaves Filho J.:
\newblock {\em { Verification of Delta Form Realization in Fixed-Point Digital Controllers Using Bounded Model Checking.}} 
\newblock In: Design Automation for Embedded Systems (to appear), pp. 49--54, 2016.
